## Predicting Weight Lifting Technique With Personal Activity Monitors

### Introduction
Personal activity monitoring data from participants performing weightlifting was downloaded from [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har). This data included a ["training" dataset](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) containing information from accelerometers on each participant's torso (belt), forearm, upper arm, and dumbell. Participants were instructed to lift the weight correctly (A), or in one of several incorrect ways (B, C, D, E). The aim of the project is to build a prediction model that identifies the method of weight lifting from accelerometer data. This model is tested on a ["validation" dataset](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv) that contains 20 cases of weight lifting to which no method (A, B, C, D, or E) has been assigned.

### Data Processing
The training and validation datasets were read into R.
```{r read, cache=TRUE}
training <- read.csv("pml-training.csv")
validation <- read.csv("pml-testing.csv")
```
The packages required for model-building, `caret` and `randomForest`, were loaded.
```{r library, message=FALSE}
library(caret); library(randomForest)
```
The "training" data was partitioned into "train" (60%) and "test" (40%) sets for model building and cross-validation, respectively. 
```{r partition}
set.seed(21)
inTrain <- createDataPartition(training$classe, p=0.6, list=FALSE)
train1 <- training[inTrain,]
test1 <- training[-inTrain,]
```
Of the 159 potential predictor variables in the dataset, the prediction function will be based on raw data from the accelerometers only. All nonmeasurement variables are removed. Most of the summary variables (e.g. average, kurtosis, skew) exhibit lots of missing values, so summary variables are removed as well. The same transformations are applied to "train", "test", and "validation" datasets.
```{r cleaning}
# Removing nonmeasurement variables
train2 <- train1[, 8:160]
test2 <- test1[, 8:160]
validation2 <- validation[, 8:160]

# Removing summary variables
col <- colnames(train2)
elimcol <- which(grepl("skewness", col)|grepl("kurtosis", col)|grepl("max", col)|grepl("min", col)|
                     grepl("amplitude", col)|grepl("stddev", col)|grepl("avg", col)|grepl("var", col))
train3 <- train2[,-elimcol]
test3 <- test2[,-elimcol]
validation3 <- validation2[,-elimcol]
```

### Model Building
To predict weight lifting method (`classe`), a random forest method was used on the modified "train" dataset. Random forest was chosen due to its high accuracy. Trials using, for example, boosting exhibited lower accuracy (results not shown).
```{r fit, cache=TRUE}
set.seed(2112)
fitrf <- train(classe~., data=train3, method="rf")

## Determining the top 5 predictors
fin <- fitrf$finalModel
importance <- fin$importance
ranked <- order(importance, decreasing=TRUE)
top5 <- ranked[1:5]
importance[top5,]
```
Based on importance, the top 5 predictors include the amount of roll and yaw detected by the belt accelerometer, the pitch detected by the forearm accelerometer, and magnet measurements in the y and z planes from the accelerometer on the dumbell.

### Cross Validation
Prior to attempting to predict weight lifting methods in the "validation" dataset, the model is first assessed on the modified "test" dataset. The random forest model is used to predict weight lifting methods for participants in the "test" dataset. These are then compared to the known weight lifting methods through calculating the confusion matrix below. 
```{r crossvalidate}
predrf <- predict(fitrf, test3)
summary <- confusionMatrix(predrf, test3$classe)
summary$table
accuracy <- summary$overall[1]
accuracy
```
Most cases of each method (A, B, C, D, and E) are predicted well by the model. The overall accuracy of the model is `r accuracy`.  

### Prediction of Unknowns
Weight lifting methods are predicted for the 20 cases in the "validation" dataset. These predictions are expected to have a lower accuracy than `r accuracy`, as out-of-sample error is usually worse than the in-sample error calculated above. However, the predictions displayed below are 100% correct (i.e. accuracy = 1), based on successful submission of these predictions in the Coursera Machine Learning class.
```{r predict}
predvalrf <- predict(fitrf, validation3)
predvalrf
```

### Conclusions
The random forest model built above accurately predicts the method of weight lifting based on measurements from accelerometers on a person's body (especially the belt and forearm) and dumbbell. This suggests that more than one personal activity monitoring device must be used if weight lifters wish to automatically assess whether their technique is correct or not. 

